Reproducible Research - Peer Assessment 1
========================================================
Author: Leslie McIntosh  
R Version: 3.1.1  
R Packages Required:
- ggplot2
- plyr
- Hmisc
- rrcovNA

```{r echotrue, echo=FALSE}
opts_chunk$set(echo=TRUE)
```


### Loading and preprocessing the data

Load the data into R. Note that I unzipped the data file before I imported it.

```{r loaddata}
activity <- read.table("~/RepData_PeerAssessment1/activity.csv", header=TRUE, sep = ",")
#Convert date formatting
activity$date <- as.Date(activity$date)
#Load required r packages
library(ggplot2)
library(plyr)
library(Hmisc)
library(plyr)
library(rrcovNA)

```
    
        
            
--------
### What is the mean total number of steps taken per day?

Make a histogram of the total number of steps taken each day

```{r stepsperday}
#library(ggplot2)
q <- qplot(date, data=activity, weight=steps, geom="histogram")
q + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Calculate and report the mean and median total number of steps taken per day

```{r meanmedian}
#library(plyr)
ddply(activity, "date", summarise, mean=mean(steps), median=median(steps))
```
--------



### What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
and the average number of steps taken, averaged across all days (y-axis)

```{r timeseries}
#library(Hmisc)
ggplot(activity,aes(x=interval,y=steps)) +
        stat_summary(fun.data ="mean_sdl", mult=1, geom = "smooth")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
(Works fine in R but not in RMarkdown)

```{r maxavgsteps}
#library(plyr)
avg_interval <- ddply(activity, .(interval), summarize,
      mean = mean(steps, na.rm=TRUE))
avg_interval[which(avg_interval$mean == max(avg_interval$mean)), ]

```
--------



### Imputing Missing Values

Calculate and report the total number of missing values in the dataset 
(i.e. the total number of rows with NAs)

```{r missingvalues}
sum(!complete.cases(activity))
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Create a new dataset that is equal to the original dataset but with the missing data filled in.

        I used the rrcovNA package for imputation as described in
        http://www.unt.edu/rss/class/Jon/Benchmarks/MissingValueImputation_JDS_Nov2010.pdf
        
```{r imputation}
#library(rrcovNA) 

summary(activity)
activity_imp <- impSeq(activity)
activity_imp <- data.frame(activity_imp)
activity_imp$date <- as.Date(activity_imp$date, origin="1970-01-01")
summary(activity_imp)
```


Make a histogram of the total number of steps taken each day.
Calculate and report the mean and median total number of steps taken per day. 
Do these values differ from the estimates from the first part of the assignment? 
What is the impact of imputing missing data on the estimates of the total daily
number of steps?

```{r stepsperday_imp}
#library(ggplot2)
q2 <- qplot(date, data=activity_imp, weight=steps, geom="histogram")
q2 + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Calculate and report the mean and median total number of steps taken per day

```{r meanmedian_imp}
#library(plyr)
ddply(activity_imp, "date", summarise, mean=mean(steps), median=median(steps))
```
--------


### Assess weekend influence

Create a new factor variable in the dataset with two levels – “weekday”
and “weekend” indicating whether a given date is a weekday or weekend day.

```{r weekend}
activity_imp$weekdays <- weekdays(activity_imp$date)

weekend_l <- c("Saturday", "Sunday")
activity_imp$weekend <- ifelse(activity_imp$weekdays %in% weekend_l, 
                        c("weekend"), c("weekday"))
table(activity_imp$weekdays, activity_imp$weekend)
```

Make a panel plot containing a time series plot (i.e. type = "l") of the
5-minute interval (x-axis) and the average number of steps taken, averaged
across all weekday days or weekend days (y-axis).

```{r panelweekend}
ggplot(activity_imp,aes(x=interval,y=steps)) +
        stat_summary(fun.data ="mean_sdl", mult=1, geom = "smooth") +
        facet_wrap(~ weekend, ncol=1)
```
